Fully Automatic Loop-free Program Synthesis by Bounded Model Checking

### 2013-10-11 paper/paper.tex
Acknowledgement

\subsection{Program Length: $l$}
The first paramter we introduce is program length, denoted by $l$.
At each iteration we synthesise programs of length exactly $l$.
We start with $l = 1$ and increment $l$ by $1$ when we have determined
that no program of length $l$ can satisfy the specification.  When we do
successfully synthesise a program, we are guaranteed that it
is of minimal length since we have previously established that no
correct program of smaller length exists.

\subsection{Word Width: $w$}
An $\mathcal{L}$-program runs on a virtual machine (the $\mathcal{L}$-machine) that
has its own set of parameters.  The only relevant parameter is
the \emph{word width} of the $\mathcal{L}$-machine, that is, the number of bits
in each internal register and immediate constant.  This parameter is denoted by
$w$.  The size of the final SAT problem generated by {\sc cbmc} scales
polynomial with $w$, since each intermediate C variable corresponds
to $w$ propositional variables.

@@

To demonstrate the capabilities of \newC as an implementation language, we synthesise a
non-trivial floating point program.  For this we chose the classic {\sc 2Sum} algorithm for computing
exact sums of floating point numbers.  The original {\sc 2Sum} algorithm is shown in Alg.~\ref{alg:2sum}.
In the case that $a \ge b$, exact addition can be implemented in fewer instructions -- this is the
{\sc Fast2Sum} algorithm~\cite{fast2sum}.  We used {\sc 2sum} as our specification, and
added an assumption that $a \ge b$.  {\sc Kalashnikov} is able to synthesise and verify the
code for {\sc Fast2Sum} from this specification in 396.07\,s, which we believe to be the first time
this has been achieved.  This provides an alternative proof to~\cite{fast2sum} that {\sc Fast2Sum} is
minimal in terms of floating point operations.

@@

\section{Conclusion}

By expressing the program synthesis problem as a safety property for a
program interpreter, we have been able to harness the power of
state-of-the-art program analysis tools and reuse them in a new problem
domain.  We have implemented our algorithm as a freely downloadable tool
whose performance compares favourably to a recent program synthesiser. 
Finally, we have taken advantage of the expressiveness of our specification
language to make an initial step towards practical synthesis of
floating-point programs.

\paragraph{Future Work}

It may be possible to extend the language $\mathcal{L}$ with constructs such
as branching and limited looping while still retaining decidability, which
would be an interesting direction to pursue.  Since our specification
language is a subset of C, it is straight-forward to interface {\sc
kalashnikov} with other analysis tools.  Various problems such as ranking
function synthesis, loop invariant generation and loop summarisation can be
phrased as program synthesis problems, and we would like to investigate the
possibilities of applying {\sc kalashnikov} to these domains.

>>>>

\subsection{Program Length: $l$}
The first paramter we introduce is program length, denoted by $l$.
At each iteration we synthesise programs of length exactly $l$.
We start with $l = 1$ and increment $l$ by $1$ when we have determined
that no program of length $l$ can satisfy the specifcation.  When we do
successfully synthesise a program, we are guaranteed that it
is of minimal length since we have previously established that no
correct program of smaller length exists.

\subsection{Word Width: $w$}
An $\mathcal{L}$-program runs on a virtual machine (the $\mathcal{L}$-machine) that
has its own set of paramters.  The only relevant parameter is
the \emph{word width} of the $\mathcal{L}$-machine, that is, the number of bits
in each internal register and immediate constant.  This parameter is denoted by
$w$.  The size of the final SAT problem generated by {\sc cbmc} scales
polynomial with $w$, since each intermediate C variable corresponds
to $w$ propositional variables.

@@

To demonstrate the capabilities of \newC as an impementation language, we synthesise a
non-trivial floating point program.  For this we chose the classic {\sc 2Sum} algorithm for computing
exact sums of floating point numbers.  The original {\sc 2Sum} algorithm is shown in Alg.~\ref{alg:2sum}.
In the case that $a \ge b$, exact addition can be implemented in fewer instructions -- this is the
{\sc Fast2Sum} algorithm~\cite{fast2sum}.  We used {\sc 2sum} as our specification, and
added an assumption that $a \ge b$.  {\sc Kalashnikov} is able to synthesise and verify the
code for {\sc Fast2Sum} from this specification in 396.07\,s, which we believe to be the first time
this has been achieved.  This provides an alternative proof to~\cite{fast2sum} that {\sc Fast2Sum} is
minimal in terms of floating point operations.

@@

\section{Conclusion}
By expressing the program synthesis problem as a safety property for a program interpreter,
we have been able to harness the power of state of the art program analysis tools and reuse
them in a new problem domain.  We have implemented our algorithm as a freely downloadable
tool whose performance compares favourably to a state of the art program synthesiser.
Fianlly, we have taken advantage of the expressiveness of our specification language
to make an initial step towards practical synthesis of floating point programs.

\paragraph{Future Work}
It may be possible to extend the language $\mathcal{L}$ with constructs such as branching and limited
looping while still retaining decidability, which would be an interesting direction to pursue.
Since our specification language is a subset of C, interfacing {\sc kalashnikov} with
other analysis tools should not be too difficult.  Various problems such as ranking function synthesis,
loop invariant generation and loop summarisation can be phrased as program synthesis problems,
and we would like to investigate the possiblities of applying {\sc kalashnikov} to these domains.

\paragraph{Acknowledgements}
We thank Martin Brain for his insight and advice during our synthesis discussions.
